---
title: "Mosaic loss of the Y chromosome estimation"
author: "Valeriia Timonina"
date: "2024-10-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
require(ggExtra)
```


## Load the data

Load the file with aggregated results for the normalised chrY coverage (result of 5_aggregate_results.sh)


```{r}
loy_data <- read.table("", header=T, sep="\t") #set a path here

head(loy_data)
```


## set a threshold to define binary mLOY trait as Q1-1.5*IQR


```{r}
loy_data$chrY_norm_scaled <- loy_data$chrY_norm - median(loy_data$chrY_norm) + 0.5 # rescale
threshold <- quantile(loy_data$chrY_norm_scaled, c(.25))-1.5*IQR(loy_data$chrY_norm_scaled) # set a threshold to define binary mLOY

loy_data$mLOY_bin <- ifelse(loy_data$chrY_norm_scaled < threshold, 1, 0) # everyone with chrY coverage below threshold are mLOY cases
```

### Visualize 

```{r}
pl <- ggplot(loy_data, aes(y=chrY_norm_scaled, x=1:length((id))))+
  geom_point(size=0.5, color="grey76")+
  geom_hline(yintercept = median(loy_data$chrY_norm_scaled), size=1.5, col="darkred")+
  geom_hline(yintercept = threshold, col="black", linetype="dashed", linewidth=1.5)+
  theme_classic()
ggExtra::ggMarginal(pl, type = "histogram", margins = "y", fill="grey76")
```

## Convert scaled normalized chrY coverage to the fraction of cells with loss of chrY

We will do it using the formula:

    = 0, if covY >= 0.5 
CF{ = -2 * covY + 1, if 0 < covY < 0.5
    = 1, if covY <= 0,

where CF - fraction of cells with LOY; covY - normalized scaled coverage of the Y chromosome.

```{r}
loy_data$mLOY_CF <- -2 * loy_data$chrY_norm_scaled + 1
loy_data$mLOY_CF <- ifelse(loy_data$chrY_norm_scaled <=0 ,1, loy_data$chrY_norm_scaled)
loy_data$mLOY_CF <- ifelse(loy_data$chrY_norm_scaled >=0.5 ,0, loy_data$chrY_norm_scaled)
```


### Visualize 

```{r}
ggplot(loy_data, aes(x=chrY_norm_scaled, y=mLOY_CF))+
  geom_point(color="grey76")+
  theme_classic()
```

